---
title: "Coefficient of Variation and Trends"
author: "the free masons"
date: "13/12/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Load the data

```{r}
library(tidyverse)
```

```{r}
load(here::here("Data", "Output",
                "student.rda"))
```


# Are the trends meaningful?

We want to understand wether the reported trends in the Math, Reading, and Science score are meaningful or it would be better to consider them random fluctuations.

To do that, we consider the coefficient of variation of the mean scores.

## Australia and New Zealand

Let's focus in Australia and New Zealand:

```{r}
student_aunz <- student %>%
  filter(country_iso3c %in% c("AUS", "NZL"))
```

For each of the three sectors we compute the weighted mean, standard deviation, quantiles, and the coefficient of variation (standard deviation normalized by the mean).

```{r}
coevar <- function(df, coevar_variable){
  df %>%
  dplyr::summarise(max = max({{ coevar_variable }}, na.rm = TRUE),
            min = min({{ coevar_variable }}, na.rm = TRUE),
            q75 = Hmisc::wtd.quantile({{ coevar_variable }}, stu_wgt, normwt = TRUE)[4],
            q25 = Hmisc::wtd.quantile({{ coevar_variable }}, stu_wgt, normwt = TRUE)[2],
            mean = Hmisc::wtd.mean({{ coevar_variable }}, stu_wgt, normwt = TRUE),
            unw_mean = mean({{ coevar_variable }}, na.rm = TRUE),
            median = median({{ coevar_variable }}, na.rm = TRUE),
            sd = sqrt(Hmisc::wtd.var({{ coevar_variable }},  normwt = TRUE)),
            #sd = sd({{ coevar_variable }}, na.rm = TRUE),
            coevar = sd / mean)
}
```


```{r}
coevar_student_math <- student_aunz %>%
  group_by(year,country_iso3c) %>%
  coevar(math)

coevar_student_science <- student_aunz %>%
  group_by(year,country_iso3c)%>%
  coevar(science)

coevar_student_read <- student_aunz %>%
  group_by(year,country_iso3c) %>%
  coevar(read)
```

```{r}
coevar_student_math %>%
  select(year, mean, unw_mean) 
```



```{r}
read <- coevar_student_read %>%
  pivot_longer(cols = c(q25, q75, max, min), names_to = "Index", values_to = "Values") %>%
  mutate(Index = as_factor(Index)) %>%
  ggplot(aes(x = year, y = Values, color = Index)) +
  geom_line() +
  geom_line(aes(y = mean)) +
  geom_errorbar(aes(ymin = mean - sd, ymax = mean + sd)) +
  ylim(0, 900) +
  facet_wrap(~ country_iso3c) +
  theme_minimal()

math <- coevar_student_math %>%
  pivot_longer(cols = c(q25, q75, max, min), names_to = "Index", values_to = "Values") %>%
  mutate(Index = as_factor(Index)) %>%
  ggplot(aes(x = year, y = Values, color = Index)) +
  geom_line() +
  geom_line(aes(y = mean)) +
  geom_errorbar(aes(ymin = mean - sd, ymax = mean + sd)) +
  ylim(0, 900) +
  facet_wrap(~ country_iso3c) +
  theme_minimal()

science <- coevar_student_science %>%
  pivot_longer(cols = c(q25, q75, max, min), names_to = "Index", values_to = "Values") %>%
  mutate(Index = as_factor(Index)) %>%
  ggplot(aes(x = year, y = Values, color = Index)) +
  geom_line() +
  geom_line(aes(y = mean)) +
  geom_errorbar(aes(ymin = mean - sd, ymax = mean + sd)) +
  ylim(0, 900) +
  facet_wrap(~ country_iso3c) +
  theme_minimal()
```

```{r}
library(patchwork)
read / math / science
```


## Australia: gender

Let's focus in Australia, and filter only specified genders:

```{r}
student_au <- student %>%
  filter(country_iso3c %in% c("AUS"),
         gender %in% c(1,2))
```

For each of the three sectors we compute the weighted mean, standard deviation, quantiles, and the coefficient of variation (standard deviation normalized by the mean).

```{r}
coevar_g <- function(df, coevar_variable){
  df %>%
  dplyr::summarise(max = max({{ coevar_variable }}, na.rm = TRUE),
            min = min({{ coevar_variable }}, na.rm = TRUE),
            q75 = Hmisc::wtd.quantile({{ coevar_variable }}, stu_wgt, normwt = TRUE)[4],
            q25 = Hmisc::wtd.quantile({{ coevar_variable }}, stu_wgt, normwt = TRUE)[2],
            mean = Hmisc::wtd.mean({{ coevar_variable }}, stu_wgt, normwt = TRUE),
            median = median({{ coevar_variable }}, na.rm = TRUE),
            sd = sqrt(Hmisc::wtd.var({{ coevar_variable }},  normwt = TRUE)),
            #sd = sd({{ coevar_variable }}, na.rm = TRUE),
            coevar = sd / mean)
}
```


```{r}
coevar_student_math_gender <- student_au %>%
  group_by(year,gender) %>%
  coevar_g(math)

coevar_student_science_gender <- student_au %>%
  group_by(year,gender)%>%
  coevar_g(science)

coevar_student_read_gender <- student_au %>%
  group_by(year,gender) %>%
  coevar_g(read)
```


```{r}
read <- coevar_student_math_gender %>%
  pivot_longer(cols = c(q25, q75, max, min), names_to = "Index", values_to = "Values") %>%
  mutate(Index = as_factor(Index)) %>%
  ggplot(aes(x = year, y = Values, color = Index)) +
  geom_line() +
  geom_line(aes(y = mean)) +
  geom_errorbar(aes(ymin = mean - sd, ymax = mean + sd)) +
  ylim(0, 900) +
  facet_wrap(~ gender) +
  theme_minimal()

math <- coevar_student_math_gender %>%
  pivot_longer(cols = c(q25, q75, max, min), names_to = "Index", values_to = "Values") %>%
  mutate(Index = as_factor(Index)) %>%
  ggplot(aes(x = year, y = Values, color = Index)) +
  geom_line() +
  geom_line(aes(y = mean)) +
  geom_errorbar(aes(ymin = mean - sd, ymax = mean + sd)) +
  ylim(0, 900) +
  facet_wrap(~ gender) +
  theme_minimal()

science <- coevar_student_science_gender %>%
  pivot_longer(cols = c(q25, q75, max, min), names_to = "Index", values_to = "Values") %>%
  mutate(Index = as_factor(Index)) %>%
  ggplot(aes(x = year, y = Values, color = Index)) +
  geom_line() +
  geom_line(aes(y = mean)) +
  geom_errorbar(aes(ymin = mean - sd, ymax = mean + sd)) +
  ylim(0, 900) +
  facet_wrap(~ gender) +
  theme_minimal()
```

```{r}
read / math / science
```

# Session Info
```{r}
sessionInfo()
```

