---
title: "Binding all country code together"
author: "Kevin Wang"
date: "19 April 2026"
output: html_document
editor_options: 
  chunk_output_type: console
---

# Loading package 

```{r}
suppressPackageStartupMessages({
  library(tidyverse)
  library(here)
  library(stringi)
})
```

# Loading School SAV files 

```{r}
extract_cnt_labels <- function(file_path) {
  # Dynamically extract year from filepath
  year_val <- as.integer(stringr::str_extract(file_path, "20[0-9]{2}"))
  
  df <- haven::read_sav(file_path, col_select = c("CNT"))
  
  # Assert that the CNT column is of class haven_labelled
  stopifnot("CNT column must be of class 'haven_labelled'" = inherits(df$CNT, "haven_labelled"))
  
  lbls <- attr(df$CNT, "labels")
  
  if (is.null(lbls)) {
    return(tibble(year = integer(), country = character(), country_name = character()))
  }
  
  tibble(
    year = year_val,
    country = as.character(lbls),
    country_name = names(lbls)
  )
}

sav_files <- c(
  here("Data/Raw/2022/CY08MSP_SCH_QQQ.SAV"),
  here("Data/Raw/2018/SCH/CY07_MSU_SCH_QQQ.sav"),
  here("Data/Raw/2015/CY6_MS_CMB_SCH_QQQ.sav")
)

countrycode_sav <- purrr::map_dfr(sav_files, extract_cnt_labels)

```

```{r}
source(here("Code/process_pisa.R"))

spss_files <- tibble(
  year = c(2012, 2009, 2006, 2003, 2000),
  syntax_file = c(
    here("Data/Raw/2012/SPSS syntax to read in school questionnaire data file.txt"),
    here("Data/Raw/2009/PISA2009_SPSS_school.txt"),
    here("Data/Raw/2006/PISA2006_SPSS_school.txt"),
    here("Data/Raw/2003/PISA2003_SPSS_school.txt"),
    here("Data/Raw/2000/PISA2000_SPSS_school_questionnaire.txt")
  ),
  data_file = c(
    here("Data/Raw/2012/INT_SCQ12_DEC03.txt"),
    here("Data/Raw/2009/INT_SCQ09_Dec11.txt"),
    here("Data/Raw/2006/INT_Sch06_Dec07.txt"),
    here("Data/Raw/2003/INT_schi_2003.txt"),
    here("Data/Raw/2000/intscho.txt")
  )
)

extract_spss_cnt_fwf <- function(year_val, syntax_file, data_file) {
  widths <- parse_spss_syntax(syntax_file)
  
  df <- readr::read_fwf(
    file = data_file, 
    col_positions = readr::fwf_positions(
      start = widths$start,
      end = widths$end,
      col_names = as.character(widths$names)
    ),
    show_col_types = FALSE
  )
  
  # Try to grab CNT if it exists, otherwise COUNTRY
  if ("CNT" %in% names(df)) {
    cnt <- as.character(df$CNT)
  } else if ("COUNTRY" %in% names(df)) {
    cnt <- as.character(df$COUNTRY)
  } else {
    return(tibble(year = integer(), country = character()))
  }
  
  # Clean NA and grab unique codes for this file/year
  cnt <- unique(cnt[!is.na(cnt)])
  
  res <- tibble(year = as.integer(year_val), country = cnt)
  return(res)
}

countrycode_spss <- purrr::pmap_dfr(
  list(spss_files$year, spss_files$syntax_file, spss_files$data_file), 
  extract_spss_cnt_fwf
)

# Manually map specific historical and transitional codes as an internal dictionary
country_manual_dict <- tibble(
  country = c("ROM", "YUG", "KHM", "SLV", "PSE", "GTM", "JAM", "MNG", "PRY", "QUR", "UZB", "TAP", "AZB", "AZE", "KGZ", "LIE", "MUS", "QCN", "QHP", "QRS", "QTN", "QVE"), 
  manual_name = c("Romania", "Yugoslavia", "Cambodia", "El Salvador", "Palestine", "Guatemala", "Jamaica", "Mongolia", "Paraguay", "Ukrainian regions (18 of 27)", "Uzbekistan", "Chinese Taipei", "Azerbaijan", "Azerbaijan", "Kyrgyzstan", "Liechtenstein", "Mauritius", "Shanghai (China)", "Himachal Pradesh (India)", "Russia (Regions)", "Tamil Nadu (India)", "Miranda (Venezuela)")
)

# Extract known valid mappings from SAV processing (use earliest known name as historical fallback)
sav_names_fallback <- countrycode_sav |> 
  dplyr::filter(!is.na(country_name) & nchar(country_name) > 0) |> 
  dplyr::group_by(country) |> 
  dplyr::slice_min(year, n = 1, with_ties = FALSE) |> 
  dplyr::ungroup() |> 
  dplyr::select(country, fallback_name = country_name)

# Map SPSS codes safely using the dictionaries
countrycode_spss_mapped <- countrycode_spss |> 
  dplyr::left_join(sav_names_fallback, by = "country") |> 
  dplyr::left_join(country_manual_dict, by = "country") |> 
  dplyr::mutate(country_name = dplyr::coalesce(manual_name, fallback_name)) |> 
  dplyr::select(year, country, country_name)

# Construct final chronological countrycode ledger
countrycode <- dplyr::bind_rows(
  countrycode_sav,
  countrycode_spss_mapped
) |> 
  dplyr::mutate(
    country_name = stringi::stri_trans_general(country_name, "Latin-ASCII"),
    country_name = dplyr::if_else(country == "TUR", "Turkey", country_name)
  ) |> 
  dplyr::arrange(desc(year), country)
```

# Historical and Supplemental Region Mappings

According to [Wikipedia's Transitional Reservations](https://en.wikipedia.org/wiki/ISO_3166-1_alpha-3#Transitional_reservations), standard country codes periodically change or omit transitional/regional states historically present in the PISA surveys. For instance:

+ `ROM` represents Romania before 2002
+ `YUG` represents Yugoslavia before 2003
+ `KHM` represents Cambodia
+ `SLV` represents El Salvador
+ `PSE` represents Palestine
+ `GTM` represents Guatemala
+ `JAM` represents Jamaica
+ `MNG` represents Mongolia
+ `PRY` represents Paraguay
+ `QUR` represents Ukrainian regions (18 of 27)
+ `UZB` represents Uzbekistan

Because standard dynamic packages often fail to capture these transitional periods accurately, we process them by mapping these specific country iso3c codes statically via a manually curated internal dictionary (`country_manual_dict`) incorporated seamlessly into the extraction pipeline above.


# Saving the Curated Country Registry

We finally export the standardized registry mapping for deployment in the main package. Note that we ensure `distinct()` is applied right before export and the chronological `year` anchor is intentionally dropped to finalize the schema.

```{r}
# Drop metadata explicitly before finalize, ensuring strict 1-to-1 mapping per ISO code,
# preferentially selecting the longest string length for country_name.
countrycode <- countrycode |> 
  dplyr::filter(!is.na(country_name) & nchar(country_name) > 0) |> 
  dplyr::mutate(
    country_name = stringi::stri_trans_general(country_name, "Latin-ASCII"),
    country_name = dplyr::if_else(
      country == "TUR", 
      "Turkey", 
      country_name
    )
  ) |> 
  dplyr::arrange(country, desc(nchar(country_name)), desc(year)) |> 
  dplyr::group_by(country) |> 
  dplyr::slice(1) |> 
  dplyr::ungroup() |> 
  dplyr::select(country, country_name) |> 
  dplyr::distinct() |> 
  dplyr::arrange(country)

full_data_transfer_folder <- "Data/Output/Transfer/data"
dir.create(here(full_data_transfer_folder), recursive = TRUE, showWarnings = FALSE)
save(countrycode, file = here(full_data_transfer_folder, "countrycode.rda"))
```

# Session Info

```{r}
sessionInfo()
```