---
title: "Binding all country code together"
author: "The Freemasons"
date: "Initiated on 19 May 2020"
output: html_document
---

# Loading package 
```{r}
library(tidyverse)
library(here)
```

# Loading data
```{r}
load(here("Data/Output/countrycode_kim.rda"))
countrycode %>% glimpse()
```

# Adding extra rows of data

According to https://en.wikipedia.org/wiki/ISO_3166-1_alpha-3#Transitional_reservations, 
+ ROM represents Romania before 2002 and 
+ YUG represents Yugoslavia before 2003. 
+ KHM represents Cambodia and
+ SLV represents El Salvador and
+ PSE represents Palestine and 
+ GTM represents Guatemala and
+ JAM represents Jamaica and
+ MNG represents Mongolia and
+ PRY represents Paraguay and
+ QUR represents Ukrainian regions (18 of 27), see https://www.oecd.org/en/publications/pisa-2022-results-volume-i-and-ii-country-notes_ed6fbcc5-en/ukrainian-regions-18-of-27_78043794-en.html#:~:text=The%20designation%20%E2%80%9CUkrainian%20regions%20(18%20of%2027)%E2%80%9D
+ UZB represents Uzbekistan.


Above countries are part of the PISA data, but sometimes standard country code names will not capture these countries. Hence, we will manually add them to what Kim has already curated. 

```{r}
countrycode <- countrycode  |>  
  bind_rows(
    tibble(
      country_iso3c = c("ROM", "YUG", "KHM", "SLV", "PSE", "GTM", "JAM", "MNG", "PRY", "QUR", "UZB"), 
      country_name = c("Romania", "Yugoslavia", "Cambodia", "El Salvador", "Palestine", "Guatemala", "Jamaica", "Mongolia", "Paraguay", "Ukrainian regions (18 of 27)", "Uzbekistan"))) |> 
  rename(country = country_iso3c)
```

```{r}
full_data_transfer_folder = "Data/Output/Transfer/data"
dir.create(here(full_data_transfer_folder), recursive = TRUE, showWarnings = FALSE)
save(countrycode, file = here(full_data_transfer_folder, "countrycode.rda"), compress = "xz")
```

# Session Info

```{r}
sessionInfo()
```