---
title: "Binding all country code together"
author: "Kevin Wang"
date: "19 April 2026"
output: html_document
editor_options: 
  chunk_output_type: console
---

# Loading package 

```{r}
suppressPackageStartupMessages({
  library(tidyverse)
  library(here)
})
```

# Loading School SAV files 

```{r}
extract_cnt_labels <- function(file_path) {
  # Dynamically extract year from filepath
  year_val <- as.integer(stringr::str_extract(file_path, "20[0-9]{2}"))
  
  df <- haven::read_sav(file_path, col_select = c("CNT"))
  
  # Assert that the CNT column is of class haven_labelled
  stopifnot("CNT column must be of class 'haven_labelled'" = inherits(df$CNT, "haven_labelled"))
  
  lbls <- attr(df$CNT, "labels")
  
  if (is.null(lbls)) {
    return(tibble(year = integer(), country = character(), country_name = character()))
  }
  
  tibble(
    year = year_val,
    country = as.character(lbls),
    country_name = names(lbls)
  )
}

sav_files <- c(
  here("Data/Raw/2022/CY08MSP_SCH_QQQ.SAV"),
  here("Data/Raw/2018/SCH/CY07_MSU_SCH_QQQ.sav"),
  here("Data/Raw/2015/CY6_MS_CMB_SCH_QQQ.sav")
)

countrycode_sav <- purrr::map_dfr(sav_files, extract_cnt_labels)

```

```{r}
source(here("Code/process_pisa.R"))

spss_files <- tibble(
  year = c(2012, 2009, 2006, 2003, 2000),
  syntax_file = c(
    here("Data/Raw/2012/SPSS syntax to read in school questionnaire data file.txt"),
    here("Data/Raw/2009/PISA2009_SPSS_school.txt"),
    here("Data/Raw/2006/PISA2006_SPSS_school.txt"),
    here("Data/Raw/2003/PISA2003_SPSS_school.txt"),
    here("Data/Raw/2000/PISA2000_SPSS_school_questionnaire.txt")
  ),
  data_file = c(
    here("Data/Raw/2012/INT_SCQ12_DEC03.txt"),
    here("Data/Raw/2009/INT_SCQ09_Dec11.txt"),
    here("Data/Raw/2006/INT_Sch06_Dec07.txt"),
    here("Data/Raw/2003/INT_schi_2003.txt"),
    here("Data/Raw/2000/intscho.txt")
  )
)

extract_spss_cnt_fwf <- function(year_val, syntax_file, data_file) {
  widths <- parse_spss_syntax(syntax_file)
  
  df <- readr::read_fwf(
    file = data_file, 
    col_positions = readr::fwf_positions(
      start = widths$start,
      end = widths$end,
      col_names = as.character(widths$names)
    ),
    show_col_types = FALSE
  )
  
  # Try to grab CNT if it exists, otherwise COUNTRY
  if ("CNT" %in% names(df)) {
    cnt <- as.character(df$CNT)
  } else if ("COUNTRY" %in% names(df)) {
    cnt <- as.character(df$COUNTRY)
  } else {
    return(tibble(year = integer(), country = character()))
  }
  
  # Clean NA and grab unique codes for this file/year
  cnt <- unique(cnt[!is.na(cnt)])
  
  res <- tibble(year = as.integer(year_val), country = cnt)
  return(res)
}

countrycode_spss <- purrr::pmap_dfr(
  list(spss_files$year, spss_files$syntax_file, spss_files$data_file), 
  extract_spss_cnt_fwf
)

# Manually map specific historical and transitional codes as an internal dictionary
country_manual_dict <- tibble(
  country = c("ROM", "YUG", "KHM", "SLV", "PSE", "GTM", "JAM", "MNG", "PRY", "QUR", "UZB", "TAP", "AZB"), 
  manual_name = c("Romania", "Yugoslavia", "Cambodia", "El Salvador", "Palestine", "Guatemala", "Jamaica", "Mongolia", "Paraguay", "Ukrainian regions (18 of 27)", "Uzbekistan", "Chinese Taipei", "Azerbaijan")
)

# Extract known valid mappings from SAV processing (use earliest known name as historical fallback)
sav_names_fallback <- countrycode_sav |> 
  dplyr::filter(!is.na(country_name) & nchar(country_name) > 0) |> 
  dplyr::group_by(country) |> 
  dplyr::slice_min(year, n = 1, with_ties = FALSE) |> 
  dplyr::ungroup() |> 
  dplyr::select(country, fallback_name = country_name)

# Map SPSS codes safely using the dictionaries
countrycode_spss_mapped <- countrycode_spss |> 
  dplyr::left_join(sav_names_fallback, by = "country") |> 
  dplyr::left_join(country_manual_dict, by = "country") |> 
  dplyr::mutate(country_name = dplyr::coalesce(manual_name, fallback_name)) |> 
  dplyr::select(year, country, country_name)

# Construct final chronological countrycode ledger
countrycode <- dplyr::bind_rows(
  countrycode_sav,
  countrycode_spss_mapped
) |> 
  dplyr::arrange(desc(year), country)
```

# Historical and Supplemental Region Mappings

According to [Wikipedia's Transitional Reservations](https://en.wikipedia.org/wiki/ISO_3166-1_alpha-3#Transitional_reservations), standard country codes periodically change or omit transitional/regional states historically present in the PISA surveys. For instance:

+ `ROM` represents Romania before 2002
+ `YUG` represents Yugoslavia before 2003
+ `KHM` represents Cambodia
+ `SLV` represents El Salvador
+ `PSE` represents Palestine
+ `GTM` represents Guatemala
+ `JAM` represents Jamaica
+ `MNG` represents Mongolia
+ `PRY` represents Paraguay
+ `QUR` represents Ukrainian regions (18 of 27)
+ `UZB` represents Uzbekistan

Because standard dynamic packages often fail to capture these transitional periods accurately, we process them by mapping these specific country iso3c codes statically via a manually curated internal dictionary (`country_manual_dict`) incorporated seamlessly into the extraction pipeline above.

# Audit Against Legacy Package Data

Before saving, we verify exactly what has changed relative to the historical dictionary previously published in the `learningtower` package. This explicit two-sided `anti_join` on both the ISO code and the region name highlights any net new entries, deprecated codes, or nomenclature adjustments.

Specifically, for nomenclature changes, the discrepancy tracking table explicitly writes out when a nomenclature shift occurred and identifies exactly which years that specific change was applied to (e.g., documenting that Turkey shifted to "Türkiye", Moldova to "Republic of Moldova", and Vietnam to "Viet Nam" in the 2022 dataset).
```{r}
# Safely load the legacy dataset without overwriting our newly built `countrycode` object
env_legacy <- new.env()
data(countrycode, package = "learningtower", envir = env_legacy)
legacy_countrycode <- env_legacy$countrycode

# Synthesize discrepancies into a unified tracking table natively
discrepancies <- dplyr::full_join(
  countrycode |> 
    dplyr::group_by(country, country_name) |> 
    dplyr::summarize(years = paste(sort(unique(year)), collapse = ", "), .groups = "drop"),
  legacy_countrycode |> dplyr::rename(legacy_name = country_name),
  by = "country"
) |> 
  dplyr::group_by(country) |> 
  dplyr::summarize(
    legacy_name = unique(legacy_name)[1],
    legacy_years = paste(na.omit(years[!is.na(country_name) & !is.na(legacy_name) & country_name == legacy_name]), collapse = " ; "),
    current_name = paste(na.omit(country_name[is.na(legacy_name) | is.na(country_name) | country_name != legacy_name]), collapse = " ; "),
    current_years = paste(na.omit(years[is.na(legacy_name) | is.na(country_name) | country_name != legacy_name]), collapse = " ; "),
    .groups = "drop"
  ) |> 
  # Isolate only the rows that do not align perfectly (i.e., any current name exists, or it was completely dropped)
  dplyr::filter(current_name != "" | (legacy_years == "" & !is.na(legacy_name))) |> 
  dplyr::mutate(
    status = dplyr::case_when(
      is.na(legacy_name) ~ "New Addition",
      current_name == "" ~ "Dropped Legacy",
      TRUE               ~ "Nomenclature Change"
    )
  ) |> 
  dplyr::arrange(status, country)

# Render beautifully using gt to explicitly classify discrepancy semantics
discrepancies |> 
  gt::gt() |> 
  gt::tab_style(
    style = gt::cell_fill(color = "#e6ffe6"), # Light green
    locations = gt::cells_body(rows = status == "New Addition")
  ) |> 
  gt::tab_style(
    style = gt::cell_fill(color = "#ffe6e6"), # Light pink
    locations = gt::cells_body(rows = status == "Dropped Legacy")
  ) |> 
  gt::tab_style(
    style = gt::cell_fill(color = "#fffbe6"), # Light yellow
    locations = gt::cells_body(rows = status == "Nomenclature Change")
  ) |> 
  gt::tab_header(
    title = "Country Nomenclature Discrepancies",
    subtitle = "Comparing freshly parsed mappings vs learningtower legacy data"
  ) |> 
  gt::cols_label(
    country = "ISO Code",
    legacy_name = "Legacy Name",
    legacy_years = "Legacy Years",
    current_name = "Current Name",
    current_years = "Current Years",
    status = "Delta"
  )
```

# Saving the Curated Country Registry

We finally export the standardized registry mapping for deployment in the main package. Note that we ensure `distinct()` is applied right before export and the chronological `year` anchor is intentionally dropped to finalize the schema.

```{r}
# Drop metadata explicitly before finalize 
countrycode <- countrycode |> 
  dplyr::select(country, country_name) |> 
  dplyr::distinct()

full_data_transfer_folder <- "Data/Output/Transfer/data"
dir.create(here(full_data_transfer_folder), recursive = TRUE, showWarnings = FALSE)
save(countrycode, file = here(full_data_transfer_folder, "countrycode.rda"), compress = "xz")
```

# Session Info

```{r}
sessionInfo()
```