---
title: "Student Quality Checks before binding"
author: "The Freemasons"
date: "`r paste0('Initiated on 2019 Dec 13, compiled on ', format(Sys.time(), '%Y %b %d'))`"
output:
  html_document:
    code_folding: hide
    fig_height: 10
    fig_width: 10
    toc: yes
    number_sections: true
    toc_depth: 3
    toc_float: yes
editor_options: 
  chunk_output_type: console
---

# Loading package 
```{r}
library(tidyverse)
library(glue)
library(here)
```

# Loading data
```{r}
## Reading in the all student data as tibbles and make sure all variables are coded in the correct variable type
dfs_stu_qqq <- tibble(years = seq(2000, 2018, 3)) %>%
  mutate(paths = glue("Data/Output/{years}/{what}.rds", what = "stu_qqq")) %>%
  mutate(data = map(.x = paths,
                    .f = ~ .x %>%
                      here() %>%
                      read_rds() %>%
                      mutate(school_id = as.character(school_id),
                             student_id = as.character(student_id)) %>%
                      mutate_at(c("mother_educ", "father_educ",
                                  "gender", "computer", "internet",
                                  "desk", "room", "dishwasher", "television",
                                  "computer_n", "car", "book"),
                                as.integer) %>%
                      mutate_at(c("math", "science", "read",
                                  "stu_wgt", "wealth", "escs"),
                                as.numeric)))

names(dfs_stu_qqq$data) <- seq(2000, 2018, 3)
student <- bind_rows(dfs_stu_qqq$data, .id = "year")

student %>%
  group_by(year) %>%
  sample_n(2)


load(here("Data/Output/countrycode.rda"))

student = student %>% 
  mutate(year = as.integer(year),
         country = as.factor(country_iso3c),
         student_id = as.factor(student_id),
         school_id = as.factor(school_id),
         gender = as.factor(gender)) %>% 
  dplyr::select(-country_iso3c) ## We will use the 3-character codes as the only identifier (factor)
```


# Quality check for each variable

We will use `janitor::tabyl` for it has a better print method and always includes NA

## Skim through data
```{r}
library(skimr)
colnames(student)
skimr::skim(student)
```


## `year`

This looks ok, though you can see how the survey was expanded in the later years.

```{r year}
library(janitor)
student %>% 
  tabyl(year)
```


## country

This looks ok.
```{r country}
student %>% 
  tabyl(country)
```

## School id

This looks ok. Though we can see that a school_id is only unique given a year country combination. So we can't track a school's performance over the years.

```{r schoolid}
student %>% 
  tabyl(school_id) %>% 
  head

student %>% 
  filter(school_id == "1") %>% 
  tabyl(country, year) %>% head(20)
```

## mother_educ

It looks like encoding for this variable changed throughout the survey. 
+ KW can confirm (through provided script in https://github.com/gvdr/learningtower_masonry/blob/master/Code/2000/student2000.R) that this variable is missing in the 2000 survey.

```{r}
student %>% 
  tabyl(year, mother_educ)
```


## father_educ

It looks like encoding for this variable changed throughout the survey. 
+ KW can confirm (through provided script in https://github.com/gvdr/learningtower_masonry/blob/master/Code/2000/student2000.R) that this variable is missing in the 2000 survey.

```{r}
student %>% 
  tabyl(year, father_educ)
```

## gender

It looks like there are inconsistent coding for NA. Needs to double check this on individual year levels. 

```{r}
student %>% 
  tabyl(year, gender)
```

## computer

It looks like there are inconsistent coding for NA. Needs to double check this on individual year levels. 

```{r}
student %>% 
  tabyl(year, computer)
```


## internet

It looks like there are inconsistent coding for NA. Needs to double check this on individual year levels. 

```{r}
student %>% 
  tabyl(year, internet)
```

## math

It looks like there are lot more missing values in the 2000 data and a small amount in 2018. Otherwise all other years are complete. 

```{r}
library(ggridges)

student %>% 
  group_by(year, is.na(math)) %>% 
  tally()

student %>% 
  ggplot(aes(x = math, y = factor(year))) +
  geom_density_ridges()
```

## read

It looks like there are lot more missing values in the 2018 data. Otherwise all other years are complete. 

2006 also contains outliers (9997) that should be removed/recoded. This coding is likely an error code. 

```{r}
student %>% 
  group_by(year, is.na(read)) %>% 
  tally()

student %>% 
  ggplot(aes(x = read, y = factor(year))) +
  geom_density_ridges()

student %>% 
  dplyr::filter(read > 1000) %>%
  group_by(year, read) %>% 
  tally()


student %>% 
  dplyr::filter(read < 1000) %>%
  ggplot(aes(x = read, y = factor(year))) +
  geom_density_ridges() +
  labs(title = "Reading score for under 1000 only")
```

## science

It looks like there are lot more missing values in the 2018 data. Otherwise all other years are complete. 

```{r}
student %>% 
  group_by(year, is.na(science)) %>% 
  tally()

student %>% 
  ggplot(aes(x = science, y = factor(year))) +
  geom_density_ridges()
```


## stu_wgt

Not sure how to judge this variable. This is highly right-skewed, but these values are also plausible observed/computed values. 

```{r}
student %>% 
  group_by(year, is.na(stu_wgt)) %>% 
  tally()

student %>% 
  ggplot(aes(x = stu_wgt, y = factor(year))) +
  geom_density_ridges()


student %>% 
  dplyr::filter(stu_wgt > 1000) %>%
  group_by(year, stu_wgt) %>% 
  tally()

student %>% 
  dplyr::filter(stu_wgt < 1000) %>%
  ggplot(aes(x = stu_wgt, y = factor(year))) +
  geom_density_ridges() +
  labs(title = "stu_wgt score for under 1000 only")
```


## desk

It looks like there are inconsistent coding for NA. Needs to double check this on individual year levels. 

```{r}
student %>% 
  tabyl(year, desk)
```


## room

It looks like there are inconsistent coding for NA. Needs to double check this on individual year levels. 

```{r}
student %>% 
  tabyl(year, room)
```

## dishwasher

It looks like there are inconsistent coding for NA. Needs to double check this on individual year levels. 

```{r}
student %>% 
  tabyl(year, dishwasher)
```



## television

It looks like there are inconsistent coding for NA. Needs to double check this on individual year levels. 

```{r}
student %>% 
  tabyl(year, television)
```

## computer_n

It looks like there are inconsistent coding for NA. Needs to double check this on individual year levels. 

```{r}
student %>% 
  tabyl(year, computer_n)
```


## car

It looks like there are inconsistent coding for NA. Needs to double check this on individual year levels. 

```{r}
student %>% 
  tabyl(year, car)
```


## book

It looks like there are inconsistent coding for NA. Needs to double check this on individual year levels. 

```{r}
student %>% 
  tabyl(year, book)
```


## wealth

Outliers (997, 999, 9997, 9999) that should be removed/recoded. These codings are likely an error codes. 

The distribution is also weirdly smooth in 2018 but discretised in 2000. 

```{r}
student %>% 
  group_by(year, is.na(wealth)) %>% 
  tally()

student %>% 
  ggplot(aes(x = wealth, y = factor(year))) +
  geom_density_ridges()

student %>% 
  dplyr::filter(wealth > 5) %>%
  group_by(year, wealth) %>% 
  tally()

student %>% 
  dplyr::filter(wealth < 5) %>%
  ggplot(aes(x = wealth, y = factor(year))) +
  geom_density_ridges() +
  labs(title = "wealth score for under 5 only")
```

## escs

Outliers (997, 999, 9997, 9999) that should be removed/recoded. These codings are likely an error codes. 

```{r}
student %>% 
  group_by(year, is.na(escs)) %>% 
  tally()

student %>% 
  ggplot(aes(x = escs, y = factor(year))) +
  geom_density_ridges()

student %>% 
  dplyr::filter(escs > 5) %>%
  group_by(year, escs) %>% 
  tally()

student %>% 
  dplyr::filter(escs < 5) %>%
  ggplot(aes(x = escs, y = factor(year))) +
  geom_density_ridges() +
  labs(title = "escs score for under 5 only")
```


# Saving data
```{r}
# save(student, file = here("Data/Output/student.rda"), compress = "xz")
```

# Session Info

```{r}
sessionInfo()
```